---
layout: post
title:  "False discovery rate control"
date:   2015-05-12
categories: mediator feature
tags: featured
---

In mulitple testing experiments, the p values from a regular t-test will introduce a lot of false positive. In *R*, the function **[p.adjust](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/p.adjust.html)** provides serveral ways to adjust the p-values.

However, the speed of *Benjamini–Hochberg procedure* in **p.adjust** is not optimized. In this case, I rewrote it in *Rcpp*, which is basically *c++*. The algorithm is explained [here](https://www.urmc.rochester.edu/biostat/people/faculty/TuSite/FDR_procedure.pdf).

The cpp function of FDR control:

```{r , warning=FALSE ,message=FALSE}
library(Rcpp)
cppFunction(' NumericVector FDRcontrol(NumericVector p, double alpha){
	// the input should be ascending order p value // 
	int size = p.size();
	double threshold1, threshold2;
	NumericVector passed(size);
	int i = 0;
	while ( i < size){
		threshold1 = alpha * (i+1) / size;
		threshold2 = alpha * (i+2) / size;
		if (threshold1 > p[i]){
			if (threshold2 < p[i+1]){
				for (int j = 0; j <= i; j++){
					passed[j] = 1;
				}
			}
		} 
		i++;
	}
	return(passed);
}')
```

Now, start benchmarking,

```{r, warning=FALSE ,message=FALSE}

library(dplyr)
library(rbenchmark)
library(data.table)

set.seed(100) #Again, reporoducible

df <- data.table(p = rnorm(100000,-0.5:0.5)) %>%
		filter(p < 1 & p > -1) %>%
		mutate(p = abs(p))	%>% 
		tbl_df
df
	
# The goal is to put a 1 if FDR < alpha
# here set FDR threshold as 0.5, which is half the time is random
benchmark(df %>% arrange(p) %>% mutate(fdr = FDRcontrol(p,0.5)),
		df %>% mutate(padj = p.adjust(p,method='BH'),FDR=ifelse(padj<0.5,1,0)))

df1 = df %>% 
		arrange(p) %>% 
		mutate(fdr = FDRcontrol(p,0.5)) %>% 
		mutate(padj = p.adjust(p,method='BH'),FDR=ifelse(padj<0.5,1,0))
df1


```

Although the *Rcpp* function is faster, it does not have the flexibility to choose a thershold for FDR after adjusting the p values. 

And also, *[Rcpp](http://www.rcpp.org)* is a good way to increase the speed of R.
