---
layout: post
title:  "Character vector splitting in R"
date:   2015-05-19
categories: mediator feature
tags: featured
---

Occasionally, I would encounter a problem in *R* where I want to split a string in a character columns with the same separator. However, there's is no function in R that is capable of doing that, and it the **strsplit** function always return a list which I have to **unlist** it.

So today, I finally typed up a *Rcpp* function to ease my work and the code is as followed.


	# include <Rcpp.h>
	using namespace Rcpp;
	
	//[[Rcpp::export]]
	std::vector< std::string > string_split(std::vector< std::string > x, char sep, int num)
	{
	    int i, j, numChar, start, length, numStrings = x.size();
		std::vector< std::string > result(numStrings);
	
	    for (i = 0 ; i < numStrings; i++ )
	    {
	        numChar = x[i].length();
			std::vector< int > sepPos(1); // which character is separater
			j = 0;
	        while (j < numChar)
	        {
	            if (x[i][j] == sep)
	            {
					sepPos.push_back(j); 
					j ++;
	            }
	            else
	            {
					j ++;
	            }
	        }
			sepPos.push_back(numChar); // add last character number
			if (num==1)
			{	
				start = sepPos[num-1] ;
				length = sepPos[num]-sepPos[num-1];
			}
			else
			{
				start = sepPos[num-1] + 1;
				length = sepPos[num]-sepPos[num-1] -1;
			}
			result[i] = x[i].substr(start,length);
			}
		}
	
	    return result;
	}


The script takes three inputs:

* The character vector that is being split
* separator 
* the fragment that is desired  



Test the code:
```{r warning = F, messages =F}

library(Rcpp)
sourceCpp('~/scripts/R/Rcpp/string_split.cpp')

testVector <- rep('I~am~a~boy',10)
for (i in 1:4){
	print(string_split(testVector,'~',i))
}
```

Benchmarking:

```{r warning = F, messages =F}
library(rbenchmark)
library(ggplot2)

r_string_split <- function(x){
	sapply(x,function(y) unlist(strsplit(x,'~'))[2])
}

bm <- benchmark(string_split(testVector,'~',2),r_string_split(testVector))
bm


ggplot(data = bm,aes(x = test, y = relative)) +
		geom_bar(stat='identity') +
		theme(axis.text.x = element_text(angle=90,
										hjust = 1,
										vjust = 0.5))+
		labs(y = 'relative speed',title = 'benchmarking result')
```

The *c++* function is almost 50x faster.
