---
layout: post
title:  "Updated: False discovery rate control"
date:   2015-05-17
categories: mediator feature
tags: featured
---

A week ago, I posted a [post](http://wckdouglas.github.io/mediator/feature/2015/05/12/BH_adjust.html) attempting to speed up FDR filtering in *R* with *Rcpp*. However, that function requires manual sorting the of the p-values before throwing into the function.

As I did some research on sorting algorithms and index sorting in *c++*, I optimized my code with **std::pair** and **std::sort** in *c++*. And surprisingly (or not), this questions has been asked a lot of times on [stackoverflow](http://stackoverflow.com/questions/10287924/c-fastest-way-to-sort-a-list-of-number-and-their-index) and other websites. 
<script src="https://gist.github.com/wckdouglas/e3121c058c4fcf88cfd5.js"></script>

So testing again:

```{r, warning=FALSE,message=FALSE}
library(data.table)
library(dplyr)
library(rbenchmark)
library(Rcpp)

sourceCpp('~/scripts/R/Rcpp/FDRcontrol.cpp')

r_equivalent <- function(pv,alpha){
	padj = p.adjust(pv,method='BH')
	return(ifelse(padj < alpha,1,0))
}

dat <- fread('~/scripts/R/Rcpp/pvalues.tsv')

alpha <- 0.01
p <- dat$p



result <- dat %>% 
		mutate(r = r_equivalent(p,alpha),
				cpp = FDRcontrol(p,alpha))
result

result %>%
	filter(r!=cpp)

```

Result is the same, so benchmarking as usual,

```{r}
benchmark(r_equivalent(p,alpha),FDRcontrol(p,alpha))
```

Apparently, *R* and *matlab*/*octave* did a great job on easing the job for user-end work. Some dat-to-day functions, such as sort and return the indices, are so easily done in high-level languages. As a biologist, I have never thought about many of these questions until I start to hack the codes. And it's always the best way to understand a concept when I start coding it out. 

